---
title: "K224161834 - Nguyễn Đức Minh Tân"
author: "Nguyễn Đức Minh Tân"
date: "13-03-2024"
output:
  html_document: default
  pdf_document: default
---

# Personal Assignment "Time Series Analytics"

## Install library

```{r cars, echo=FALSE}
install.packages("fpp2")
install.packages("gglot2")
install.packages("tsibble")
```

## Run library

```{r cars, echo=FALSE}
library(fpp2)
library(ggplot2)
library(fma)
library(forecast)
library(expsmooth)
library(tsibble)
```

## Dataset 'hsales' overview

```{r pressure, echo=FALSE}
data("hsales")
summary(hsales)
str(hsales)
head(hsales)
tail(hsales)
```

```{r}
# Define the sales data
sales_data <- tibble(
  Year = rep(1973:1995, each=12),
  Month = rep(c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"), times=23),
  Sales = c(55, 60, 68, 63, 65, 61, 54, 52, 46, 42, 37, 30,
            37, 44, 55, 53, 58, 50, 48, 45, 41, 34, 30, 24,
            29, 34, 44, 54, 57, 51, 51, 53, 46, 46, 46, 39,
            41, 53, 55, 62, 55, 56, 57, 59, 58, 55, 49, 47,
            57, 68, 84, 81, 78, 74, 64, 74, 71, 63, 55, 51,
            57, 63, 75, 85, 80, 77, 68, 72, 68, 70, 53, 50,
            53, 58, 73, 72, 68, 63, 64, 68, 60, 54, 41, 35,
            43, 44, 44, 36, 44, 50, 55, 61, 50, 46, 39, 33,
            37, 40, 49, 44, 45, 38, 36, 34, 28, 29, 27, 29,
            28, 29, 36, 32, 36, 34, 31, 36, 39, 40, 39, 33,
            44, 46, 57, 59, 64, 59, 51, 50, 48, 51, 45, 48,
            52, 58, 63, 61, 59, 58, 52, 48, 53, 55, 42, 38,
            48, 55, 67, 60, 65, 65, 63, 61, 54, 52, 51, 47,
            55, 59, 89, 84, 75, 66, 57, 52, 60, 54, 48, 49,
            53, 59, 73, 72, 62, 58, 55, 56, 52, 52, 43, 37,
            43, 55, 68, 68, 64, 65, 57, 59, 54, 57, 43, 42,
            52, 51, 58, 60, 61, 58, 62, 61, 49, 51, 47, 40,
            45, 50, 58, 52, 50, 50, 46, 46, 38, 37, 34, 29,
            30, 40, 46, 46, 47, 47, 43, 46, 37, 41, 39, 36,
            48, 55, 56, 53, 52, 53, 52, 56, 51, 48, 42, 42,
            44, 50, 60, 66, 58, 59, 55, 57, 57, 56, 53, 51,
            45, 58, 74, 65, 65, 55, 52, 59, 54, 57, 45, 40,
            47, 47, 60, 58, 63, 64, 64, 63, 55, 54, 44, 40)
)
```

The 'hsales' data contains monthly sales of new one-family houses sold in the USA from 1973 to 1995.

### Visualization and exploration

```{r, echo=FALSE}
plot(hsales, main="Sales of new one-family houses in the USA", xlab="Year", ylab="Number sold")
```

We can initially see that there is some seasonality in the data. Next, we plot the decomposition of the time series into seasonal, trend, and remainder components.

```{r, echo=FALSE}
plot(stl(hsales, s.window="periodic"))
```

### Plots

#### Time plots

```{r}
autoplot(hsales)
```

```{r}
monthplot(hsales, main="Monthly plot: Sales of new one-family houses in the USA", xlab="Month", ylab="Number sold")
data('hsales')
```

#### **Transforming the time series**

```{r}
trans <- seasadj(stl(hsales, s.window = "periodic"))
tsdisplay(trans)
```

I will model the trend component graphed above hsales. I will use a seasonal ARIMA model for this, which means we must transform the data into a stationary time series. The high PACF at lag 1 suggests we should compare the data with lag 1.

```{r}
tsdisplay(diff(trans))
```

#### Histogram

```{r}
hist(hsales, main="Sales histogram of new one-family houses in the USA", xlab="Number sold", ylab="Frequency")
```

#### Boxplot

```{r}
boxplot(hsales, main="Boxplot of Sales of new one-family houses in the USA", ylab="Number sold")
```

#### **Forecasting 1**

```{r}
fit <- Arima(diff(trans), order=c(1, 0, 0))
summary(fit)
```

```{r}
fit <- Arima(diff(trans), order=c(1, 0, 1))
summary(fit)
```

```{r}
plot(forecast(fit, h=24))
```

#### Forecasting 2

```{r}
# split data in training and test sets
head.ts <- function(hsales,n) {
  hsales <- as.ts(hsales)
  window(hsales,end=tsp(hsales)[2]-(n-1)/frequency(hsales))
}
training <- head(hsales,24)
training
```

```{r}
tail.ts <- function(hsales,n) {
  hsales <- as.ts(hsales)
  window(hsales,start=tsp(hsales)[2]-(n-1)/frequency(hsales))
}
test <- tail(hsales,23)
test
```

```{r}
# Benchmark forcastng methodes
hsalesfit1 <- meanf(training, h=11)
hsalesfit2 <- naive(training, h=11)
hsalesfit3 <- snaive(training, h=11)
hsalesfit4 <- rwf(training, h=11, drift=TRUE)
plot(hsalesfit1, plot.conf=FALSE, ylab="Number Sold", xlab="Year",
     main="Forecasts for two years production")
lines(hsalesfit2$mean,col=2)
lines(hsalesfit3$mean,col=3)
lines(hsalesfit4$mean,col=4)
legend("topleft",lty=1,col=c(4,2,3,5),
      legend=c("Mean method","Naive method","Seasonal naive method", "Drift method"))
```

```{r}
# residuals for the best methode (snaive)
res <- residuals(snaive(hsales))
plot(res, main="Residuals from snaive method",
ylab="", xlab="Year")
```

## Dataset 'gasoline' overview

```{r}
data("gasoline")
summary(gasoline)
str(gasoline)
```
